---
title: "VJ Surgical Methods Meta Analyses"
output: html_notebook
---




# Read in the data

```{r}

df <- read.csv('VJ_meta_analysis_data_preprocessing.csv')
df
```


# Single Meta Analysis, Random-Effects, Inverse Variance

```{r}
df.radiofreq <- df[df['Method'] =='Radiofrequency', ]

df.radiofreq
```



## metafor escalc

To cite the package, please use the following reference:

Viechtbauer, W. (2010). Conducting meta-analyses in R with the metafor package. Journal of Statistical Software, 36(3), 1–48. https://www.jstatsoft.org/v036/i03.


Use help(escalc) and read subheading "Measures for Dichotomous Variables"
Measures for Dichotomous Variables
A meta-analysis may be conducted to aggregate studies that provide data for individual groups with respect to a dichotomous dependent variable. Here, one needs to specify xi and ni, denoting the number of individuals experiencing the event of interest and the total number of individuals, respectively. Instead of specifying ni, one can use mi to specify the number of individuals that do not experience the event of interest. The options for the measure argument are then:

"PR" for the raw proportion.

"PLN" for the log transformed proportion.

"PLO" for the logit transformed proportion (i.e., log odds).

"PAS" for the arcsine square root transformed proportion (i.e., the angular transformation).

"PFT" for the Freeman-Tukey double arcsine transformed proportion (Freeman & Tukey, 1950).

Zero cell entries can be problematic for certain outcome measures. When to="only0" (the default), the value of add (the default is 1/2; but see ‘Note’) is added to xi and mi only for studies where xi or mi is equal to 0. When to="all", the value of add is added to xi and mi in all studies. When to="if0all", the value of add is added in all studies, but only when there is at least one study with a zero value for xi or mi. Setting to="none" or add=0 has the same effect: No adjustment to the observed values is made. Depending on the outcome measure and the data, this may lead to division by zero inside of the function (when this occurs, the resulting value is recoded to NA).

```{r}

require(metafor)

help(escalc)
help(forest)
# help(rma)

```


```{r}
df.radiofreq.PR <- escalc(data=df.radiofreq, xi=SF, ni=Total, measure="PR")

df.radiofreq.PR
```

do the meta analysis using rma

```{r}

df.radiofreq.PR.rma <- rma(df.radiofreq.PR, slab=Study, method="REML", weighted=TRUE)
```

```{r}
forestplot_radiofreq <- forest(df.radiofreq.PR.rma, 
       header=c("Radiofrequency Study", "Proportion Seizure Free [95% CI]"),
       mlab="Random Effects",
       order=c(4,1,5, 2, 3, 6, 8,7),
       alim=c(0, 1), steps=6, xlim=c(-0.5, 1.5),
       )

dev.copy(device=svg, "Figures\\radiofreq.svg", width=8, height=5)
dev.off()
```
```{r}
funnel(df.radiofreq.PR.rma,
       pch.fill=21,
       )
help(funnel)
```


```{r}
funnel(trimfill(df.radiofreq.PR.rma), pch.fill = 21)
trimfill(df.radiofreq.PR.rma)
```










# Adjustment / Moderators of follow-up duration

```{r}

# read subset of data
df <- read.csv('VJ_meta_analysis_data_preprocessing.csv')
df.radiofreq <- df[df['Method'] =='Radiofrequency', ]

# effect size calculation
require(metafor)
df.radiofreq.PR <- escalc(data=df.radiofreq, xi=SF, ni=Total, measure="PR")

# random effects model and moderators
df.radiofreq.PR.rma_mods <- rma(df.radiofreq.PR, slab=Study, method="REML", weighted=TRUE,
                                mods=Avg.FollowUp)

```

```{r}
#forest plots and save
forestplot_radiofreq <- forest(df.radiofreq.PR.rma_mods, 
       header=c("Radiofrequency with Follow-up Duration as Moderator", "Seizure Freedom [95% CI]"),
       mlab="Random Effects",
       order=c(4,1,5, 2, 3, 6, 8,7),
       alim=c(0, 1), steps=6, xlim=c(-0.5, 1.5),
       )
dev.copy(device=svg, "Figures\\radiofreq_mods.svg", width=8, height=5)
dev.off()

```
Caption: 
he grey-colored diamonds (or polygons) are the fitted values and the width of the diamonds/polygons reflects the width of the CI for the fitted values.
https://stackoverflow.com/questions/57513712/does-metafor-package-in-r-provide-forest-plot-for-robust-random-effects-models




```{r}
# # funnel plots and save
# funnel(trimfill(df.radiofreq.PR.rma_mods), pch.fill = 21)
# trimfill(df.radiofreq.PR.rma_mods)
# dev.copy(device=svg, "Figures\\radiofreq_mods_trimfill.svg", width=8, height=5)
# dev.off()
```








# Subplotting all methods


```{r}

# effect size calculation
df.PR <- escalc(data=df, xi=SF, ni=Total, measure="PR")

# random effects model and moderators
df.PR.rma <- rma(df.PR, slab=Study, method="REML", weighted=TRUE,
                                # subset=df.PR$Method=="Radiofrequency",
                 )
#forest plots and save
forestplot_df <- forest(df.PR.rma, 
       header=c("Surgical Methods", "Proportion Seizure Free [95% CI]"),
       mlab="Random Effects",
       order=df.PR.rma$Method,
       # alim=c(0, 1), steps=6, xlim=c(-0.5, 1.5),
       )
dev.copy(device=svg, "Figures\\subplots_fail.svg", width=10, height=8)
dev.off()
```






https://www.metafor-project.org/doku.php/plots:forest_plot_with_subgroups



```{r, fig.width=10, fig.height=14}

# effect size calculation
df.PR <- escalc(data=df, xi=SF, ni=Total, measure="PR")

# random effects model and moderators
df.PR.rma <- rma(df.PR, slab=Study, method="REML", weighted=TRUE,
                                # subset=df.PR$Method=="Radiofrequency",
                 )






#forest plots 
forest(df.PR.rma, 
       header=c("Surgical Methods", "Proportion Seizure Free [95% CI]"),
       mlab="",  #"Random Effects",
       order=df.PR.rma$Method,
       alim=c(0, 1), steps=6, xlim=c(-1.6, 1.5), ylim=c(-1, 85),
       cex=0.7,
       )


### add text with Q-value, dfs, p-value, and I^2 statistic
overall_text <- text(-1.6, -1, pos=4, cex=0.75, bquote(paste("RE for All Studies (Q = ",
     .(formatC(df.PR.rma$QE, digits=2, format="f")), ", df = ", .(df.PR.rma$k - df.PR.rma$p),
     ", p = ", .(formatC(df.PR.rma$QEp, digits=2, format="f")), "; ", I^2, " = ",
     .(formatC(df.PR.rma$I2, digits=1, format="f")), "%)")))

### fit meta-regression model to test for subgroup differences
res <- rma(df.PR, mods = ~ Method)

### add text for the test of subgroup differences
text(-1.6, -2.4, pos=4, cex=0.75, bquote(paste("Test for Subgroup Differences: ",
     Q[M], " = ", .(formatC(res$QM, digits=2, format="f")), ", df = ", .(res$p - 1),
     ", p = ", .(formatC(res$QMp, digits=2, format="f")))))

### set font expansion factor (as in forest() above) and use bold italic
### font and save original settings in object 'op'
op <- par(cex=0.75, font=4)

### add text for the subgroups
text(-1.6, c(2.4, 1.6, 0.5), pos=4, c("Radiofrequency",
                               "ATL",
                               "Gamma Knife"))
 

### switch to bold font
par(font=2)
 
# ...


### set par back to the original settings
par(op)
 
### fit random-effects model in the three subgroups
res.rf <- rma(df.PR, subset=(Method=="Radiofrequency"))
res.atl <- rma(df.PR, subset=(Method=="ATL"))
res.gn <- rma(df.PR, subset=(Method=="Gamma Knife"))

### add summary polygons for the three subgroups
addpoly(res.rf, row=18.5, cex=0.75, atransf=exp, mlab="")
addpoly(res.atl, row= 7.5, cex=0.75, atransf=exp, mlab="")
addpoly(res.gn, row= 1.5, cex=0.75, atransf=exp, mlab="")

## and save
# dev.copy(device=svg, "Figures\\subplots_df.svg", width=10, height=8)
# dev.off()
```


check the order of the subgroups cf csv spreadsheet
```{r}
order(df.PR$Method)
```
































# Main Subplots all-in-1



therefore order is bottom to up:
ATL
Gamma Knife
LiTT
RF
transcortical
transylvian
unspecified
Surgery 


set the initial values:

```{r}
x_begin = -1.6
y_begin = -1.5
y_end = 120

ilab_x_1 = -1
ilab_x_2 = -0.8

gap =5

y_end = y_end + gap*7

fontsize = 0.25

```


do meta analysis and plot:

```{r, fig.width=10, fig.height=30}

# effect size calculation
df.PR <- escalc(data=df, xi=SF, ni=Total, measure="PR")

# random effects model and moderators
df.PR.rma <- rma(df.PR, slab=Study, method="REML", weighted=TRUE,
                                # subset=df.PR$Method=="Radiofrequency",
                 )


# - gaps of 5
# ATL 23
# Gamma 15
# LITT 14
# RF 8
# transcortical 14
# transsylvian 8 
# unspecified 2
# Surgery 1





#forest plots # rwos is bottom up
forest(df.PR.rma, 
       header=c("Surgical Techniques and Studies", "Proportion Seizure Free [95% CI]", cex=2),
       mlab="",  #"Random Effects",
       ilab=cbind(df.PR$SF, df.PR$Total),
       ilab.xpos=c(ilab_x_1, ilab_x_2),
       order=order(df.PR$Method),
       at=c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), 
       xlim=c(x_begin, 1.5), 
       refline=0.666,
       # showweights = TRUE,
       ylim=c(y_begin, y_end),
       rows=c(3:(3+22), (30+gap):(30+14+gap), (49+gap*2):(49+13+gap*2), (49+13+5+gap*3):(74+gap*3), (79+gap*4):(92+gap*4), (97+gap*5):(97+8-1+gap*5), (97+8-1+5+gap*6):(110+gap*6), (115+gap*7):(115+gap*7)),
       cex=fontsize,
       addcred=FALSE, level=95,
       )


### add text with Q-value, dfs, p-value, and I^2 statistic
text(x_begin, -2.5, pos=4, cex=0.4, bquote(paste("RE for All Studies (Q = ",
     .(formatC(df.PR.rma$QE, digits=2, format="f")), ", df = ", .(df.PR.rma$k - df.PR.rma$p),
     ", p = ", .(formatC(df.PR.rma$QEp, digits=2, format="f")), "; ", I^2, " = ",
     .(formatC(df.PR.rma$I2, digits=1, format="f")), "%)")))

### fit meta-regression model to test for subgroup differences
res <- rma(df.PR, mods = ~ Method)

### add text for the test of subgroup differences
text(x_begin, -5.5, pos=4, cex=0.4, bquote(paste("Test for Subgroup Differences: ",
     Q[M], " = ", .(formatC(res$QM, digits=2, format="f")), ", df = ", .(res$p - 1),
     ", p = ", .(formatC(res$QMp, digits=2, format="f")))))

## set font expansion factor (as in forest() above) and use bold italic
## font and save original settings in object 'op'
op <- par(cex=0.75, font=4)

### add text for the subgroups
text(x_begin, cex=0.45, c((26), (45+gap), (63+gap*2), (75+gap*3), (93+gap*4), (105+gap*5), (111+gap*6), (116+gap*7)), pos=4, c("ATL",
                                                             "Gamma Knife",
                                                             "LiTT",
                                                             "Radiofrequency",
                                                             "SAH Transcortical",
                                                             "SAH Transsylvian",
                                                             "SAH Unspecified",
                                                             "Epilepsy Surgery - Any"))


### switch to bold font
par(font=2)


### add column headings to the plot
text(c(ilab_x_1,ilab_x_2), y_end-1, c("Seizure Free (Engel I)", "Total"), cex=0.5)
# text(c(-8.75,-5.25),     27, c("Vaccinated"))


### set par back to the original settings
par(op)



### fit random-effects model in the 7 subgroups excluding surgery
# ATL
# Gamma Knife
# LiTT
# RF
# transcortical
# transylvian
# unspecified
# Surgery 
res.rf <- rma(df.PR, subset=(Method=="Radiofrequency"))
res.atl <- rma(df.PR, subset=(Method=="ATL"))
res.gn <- rma(df.PR, subset=(Method=="Gamma Knife"))
res.litt <- rma(df.PR, subset=(Method=="LiTT"))
res.tc <- rma(df.PR, subset=(Method=="SAH_transcortical"))
res.ts <- rma(df.PR, subset=(Method=="SAH_transylvian"))
res.SAH <- rma(df.PR, subset=(Method=="SAH_unspecified"))

 
# ### add summary polygons for the subgroups
addpoly(res.SAH, row=(106.5+gap*6), cex=0.3, mlab="")
addpoly(res.ts, row= (94.5+gap*5), cex=0.3, mlab="")
addpoly(res.tc, row= (76.5+gap*4), cex=0.3, mlab="")
addpoly(res.rf, row= (49+13+2.5+gap*3), cex=0.3, mlab="")
addpoly(res.litt, row= (46.5+gap*2), cex=0.3, mlab="")
addpoly(res.gn, row=(27.5+gap), cex=0.3, mlab="")
addpoly(res.atl, row= 1, cex=0.3, mlab="")


### add text with Q-value, dfs, p-value, and I^2 statistic for subgroups. 2nd number is the row as above
text(x_begin, (106.5+gap*6), pos=4, cex=fontsize, bquote(paste("RE Model for Subgroup (Q = ",
     .(formatC(res.SAH$QE, digits=2, format="f")), ", df = ", .(res.SAH$k - res.SAH$p),
     ", p = ", .(formatC(res.SAH$QEp, digits=2, format="f")), "; ", I^2, " = ",
     .(formatC(res.SAH$I2, digits=1, format="f")), "%)")))

res.s <- res.ts
text(x_begin, (94.5+gap*5), pos=4, cex=fontsize, bquote(paste("RE Model for Subgroup (Q = ",
     .(formatC(res.s$QE, digits=2, format="f")), ", df = ", .(res.s$k - res.s$p),
     ", p = ", .(formatC(res.s$QEp, digits=2, format="f")), "; ", I^2, " = ",
     .(formatC(res.s$I2, digits=1, format="f")), "%)")))

res.s <- res.tc
text(x_begin, (76.5+gap*4), pos=4, cex=fontsize, bquote(paste("RE Model for Subgroup (Q = ",
     .(formatC(res.s$QE, digits=2, format="f")), ", df = ", .(res.s$k - res.s$p),
     ", p = ", .(formatC(res.s$QEp, digits=2, format="f")), "; ", I^2, " = ",
     .(formatC(res.s$I2, digits=1, format="f")), "%)")))

res.s <- res.rf
text(x_begin, (49+13+2.5+gap*3), pos=4, cex=fontsize, bquote(paste("RE Model for Subgroup (Q = ",
     .(formatC(res.s$QE, digits=2, format="f")), ", df = ", .(res.s$k - res.s$p),
     ", p = ", .(formatC(res.s$QEp, digits=2, format="f")), "; ", I^2, " = ",
     .(formatC(res.s$I2, digits=1, format="f")), "%)")))

res.s <- res.litt
text(x_begin, (46.5+gap*2), pos=4, cex=fontsize, bquote(paste("RE Model for Subgroup (Q = ",
     .(formatC(res.s$QE, digits=2, format="f")), ", df = ", .(res.s$k - res.s$p),
     ", p = ", .(formatC(res.s$QEp, digits=2, format="f")), "; ", I^2, " = ",
     .(formatC(res.s$I2, digits=1, format="f")), "%)")))

res.s <- res.gn
text(x_begin, (27.5+gap), pos=4, cex=fontsize, bquote(paste("RE Model for Subgroup (Q = ",
     .(formatC(res.s$QE, digits=2, format="f")), ", df = ", .(res.s$k - res.s$p),
     ", p = ", .(formatC(res.s$QEp, digits=2, format="f")), "; ", I^2, " = ",
     .(formatC(res.s$I2, digits=1, format="f")), "%)")))

res.s <- res.atl
text(x_begin, 1, pos=4, cex=fontsize, bquote(paste("RE Model for Subgroup (Q = ",
     .(formatC(res.s$QE, digits=2, format="f")), ", df = ", .(res.s$k - res.s$p),
     ", p = ", .(formatC(res.s$QEp, digits=2, format="f")), "; ", I^2, " = ",
     .(formatC(res.s$I2, digits=1, format="f")), "%)")))





# and save
dev.copy(device=svg, "Figures\\MAIN_subplots_v2.svg", width=10, height=8)
dev.off()
```





```{r}

length(c(3:(3+22), 30+gap:(30+14+gap), 49+gap:(49+13+gap), (49+13+5+gap):(74+gap), (79+gap):(92+gap), (97+gap):(97+8-1+gap), (97+8-1+5+gap):(110+gap), (115+gap):(115+gap)))
```






```{bash}
git push -f
```















